Millions of Type Errors

### 2023-09-14 programming23/paper.tex
abstract draft

  \paragraph{Context}
  %% What is the broad context of the work? What is
  %% the importance of the general research area?
  {Roblox Studio} is a tool that lets millions of creators
  build interactive experiences by programming in a variant
  of Lua called Luau.
  The creators form a broad group, ranging from novices writing
  their first script to professional Luau developers, thus Luau
  must support a wide audience.
  As part of its efforts to support all kinds programmers, Luau includes a
  gradual type system that goes to great lengths to minimize false positive
  errors.

  \paragraph{Inquiry}
  %% What problem or question does the paper
  %% address? How has this problem or question been
  %% addressed by others (if at all)?
  Since Luau is in the hands of so many creators,
  how do they use it?
  Of central importance, however, is protecting the privacy
  of Luau creators.
  FILL

%  TYPE ERRORS
%  investigate

  It has the ability to report telemetry data,
  which allows large-scale measurement of the creator experience. In
  this paper, we discuss one use of telemetry, to measure the experience
  of type error reporting.
  We find FILL.

  \paragraph{Approach}
  %% What was done that unveiled new knowledge?
  FILL impersonal telemetry, N months.

  \paragraph{Knowledge}
  %% What new facts were uncovered? If the
  %% research was not results oriented, what new
  %% capabilities are enabled by the work?
  FILL

  \paragraph{Grounding}
  %% What argument, feasibility proof, artifacts,
  %% or results and evaluation support this work?
  Our findings are supported by a publicly-available dataset
  of over 1.5 million telemetry records.

  \paragraph{Importance}
  %% Why does this work matter?
  FILL improve Luau, offer lessons to similar developments (epic?, elixir?).

@@

In order to avoid swamping our telemtry servers, we randomly throttle
the generation of telemetry records:
\begin{itemize}
  \item
    10\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    5\% of runs of the typechecker in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we additionally have:
\begin{itemize}
  \item
    every run of the typechecker in an enrolled session which switched file generated a telemetry record.
\end{itemize}
Each telemetry record contains:
\begin{description}
  \item[D1] metadata: a psedonymized session identifier, timestamp and session duration,
  \item[D2] mode of the current file: (\mnocheck{}, \mnonstrict{}, or \mstrict{}),
  \item[D3] reason for sending: (random selection, or file switch),
  \item[D4] size of codebase: number of files, and number of lines of code type checked,
  \item[D5] size of edit range
  \item[D6] overall type errors: for previous and current states, we have the total number, the nunber in the current file, and number in the edit range,
  \item[D7] specific errors in the edit range: for the previous and current states, for each type error code, the number type errors in the edit range,
  \item[D8] forced strict errors: specific errors for the
    edit range computed as if the user had strict mode enabled.
  \item[D9] TooComplexErrors: project-wide count for this one specific error type
\end{description}
To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large edit
ranges (for example if the user edits an the beginning and end of the
file) but fits in a fixed-size telemetry record.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}

@@

Error counts per-project and per-module include syntax errors.
Problematic, because syntax errors are uninteresting (easy to fix),
but common, which makes it likely that our per-keystroke telemetry (1/2k) will record them.
Edit range has specific counts.
Next time, we should do that for the other aggregates.

@@

>>>>

  \paragraph{Context}
  %% What is the broad context of the work? What is
  %% the importance of the general research area?
  {Roblox Studio} lets millions of creators
  build interactive experiences by programming in a variant
  of Lua called Luau.
  The creators form a broad group, ranging from novices writing
  their first script to professional developers, thus Luau
  must support a wide audience.
  As part of its efforts to support all kinds programmers, Luau includes a
  gradual type system and goes to great lengths to minimize false positive
  errors.

  \paragraph{Inquiry}
  %% What problem or question does the paper
  %% address? How has this problem or question been
  %% addressed by others (if at all)?
  Since Luau is in the hands of many creators, we want to collect their feedback
  and ultimately improve the language design (in particular, the type system).
  We want feedback at a large scale, as creators work on their actual projects,
  which suggests the deployment of client-side telemetry.
  %% ^ awkward
  However, we must respect creators privacy.
  Whatever telemetry we use to study the Luau language cannot include
  personally-identifiable information.
  For example, we cannot monitor source code or error messages as they
  may contain personal data.
  Finding ways to harness telemetry in a private way is the central research challenge.

  \paragraph{Approach}
  %% What was done that unveiled new knowledge?
  We designed and implemented a privacy-preserving telemetry system for Luau.
  Telemetry records include a timestamp, a (randomized) session id, a reason
  for sending, and a sequence of counter values summarizing the results of type
  analysis.
  Over four months in Spring 2023, we deployed the telemetry system and collected
  1.5 million telemetry records.
  We analyzed the data to improve Luau.

  \paragraph{Knowledge}
  %% What new facts were uncovered? If the
  %% research was not results oriented, what new
  %% capabilities are enabled by the work?
  We present several findings about Luau:
  some confirm hypotheses by the Luau team, others are rather surprising, but
  all findings support our claim that privacy-preserving telemetry is a useful
  way to collect feedback.
  One expected finding is that types are unpopular; there is an order-of-maginitude
  difference in the number of sessions that use different type analysis modes.
  One surprise is that the strict type analysis mode evidently makes it difficult
  to connect a script with data assets.
  Lastly, one reassuring finding is that type analysis rarely hits its internal
  limits on problem size.

  \paragraph{Grounding}
  %% What argument, feasibility proof, artifacts,
  %% or results and evaluation support this work?
  Our findings are supported by a publicly-available dataset
  of over 1.5 million telemetry records
  and by free software for analyzing the data.

  \paragraph{Importance}
  %% Why does this work matter?
  Beyond the immediate benefits to Luau,
  our findings about types and type errors have implications
  for adoption and ergonomics in other gradually typed languages.
  %% mention Elixir, Verse?
  FILL specifics.
  Our telemetry design is of broad interest, as it gathers useful
  data without exposing personal information.

@@

In order to avoid swamping our telemtry servers, we randomly throttle
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we additionally have:
\begin{itemize}
  \item
    every run of the typechecker in an enrolled session which switched file generated a telemetry record.
\end{itemize}
Each telemetry record contains:
\begin{description}
  \item[D1] metadata: a psedonymized session identifier, timestamp and session duration,
  \item[D2] mode of the current file: (\mnocheck{}, \mnonstrict{}, or \mstrict{}),
  \item[D3] reason for sending: (random selection, or file switch),
  \item[D4] size of codebase: number of files, and number of lines of code type checked,
  \item[D5] size of edit range
  \item[D6] overall type errors: for previous and current states, we have the total number, the nunber in the current file, and number in the edit range,
  \item[D7] specific errors in the edit range: for the previous and current states, for each type error code, the number type errors in the edit range,
  \item[D8] forced strict errors: specific errors for the
    edit range computed as if the user had strict mode enabled.
  \item[D9] TooComplexErrors: project-wide count for this one specific error type
\end{description}
To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large edit
ranges (for example if the user edits an the beginning and end of the
file) but fits in a fixed-size telemetry record.

@@

  \mnocheck{}
  \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--te-density-diff.pdf}
  \medskip
  \mnonstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--te-density-diff.pdf}
  \medskip
  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}

@@

Error counts per-project and per-module include syntax errors.
Problematic, because syntax errors are uninteresting (easy to fix),
but common, which makes it likely that our per-keystroke telemetry (1/200) will record them.
Edit range has specific counts.
Next time, we should do that for the other aggregates.

@@

\paragraph{Do we need timestamps?}

Our telemetry records include timestamps.
While timestamps are not personal data per se, they are in the ballpark.
Telemetry would be even more private without them.
Do we need them?

Timestamps are not the focus of our analysis, but they do play a critical role.
Consider~\cref{f:error-delta}.
With timestamps, it is clear that error density stays close to zero.
Without timestamps, we would be left with the counts at the bottom
of~\cref{f:error-delta}.
Many distributions could fit those counts, such as a ``mountain'' session that first
grows and grows the number of type errors then shrinks and shrinks down to zero.
